---
title: "Course Project 1"
date: "Date: `r format(Sys.Date(),'%d-%m-%Y')`"
output: html_document
---

#### 1. Code for reading in the dataset and processing the data

```{r}
df <- read.csv("activity.csv", sep = ",")
str(df)
df$date <- as.Date(df$date, format= "%Y-%m-%d")
```

#### 2. Histogram of the total number of steps taken each day

```{r}
library(dplyr)
library(dplyr)
temp <- df %>% group_by(date) %>% summarise(totalstep = sum(steps, na.rm = T))
hist(temp$totalstep, xlab = "No.of steps each day", main = "Histogram of the total number of steps taken each day")

```

#### 3. Mean and median number of steps taken each day: 9354.23 & 10395, respectively
```{r}
mean(temp$totalstep)
median(temp$totalstep)

```

#### 4.Time series plot of the average number of steps taken
```{r}
temp <- df %>% group_by(interval) %>% summarise(average_step_interval = mean(steps, na.rm = T))
with(temp, plot(interval, average_step_interval, type="l", xlab="Time interval", ylab="Average steps", main="The average daily activity pattern"))

```

#### 5. The 5-minute interval that, on average, contains the maximum number of steps:835
```{r}
temp$interval[temp$average_step_interval == max(temp$average_step_interval )]

```

#### 6. Code to describe and show a strategy for imputing missing data
- The total number of missing values in the dataset (i.e. the total number of rows with NAs): 2304
```{r}
sapply(df, function(x) sum(is.na(x)))
sum(is.na(df$steps))
```

- A strategy for filling in all of the missing values in the dataset: NA values will be filled by the mean for that 5-minute interval.A new dataset that is equal to the original dataset but with the missing data filled in:
```{r}
newdf <- merge(df, temp, by = "interval", all.x = T)
newdf <- newdf %>% mutate(new_steps = ifelse(is.na(steps), average_step_interval, steps))

```

- New mean and median total number of steps taken per day: 10766.19 for both due to the imputing method
```{r}
temp <- newdf %>% group_by(date) %>% summarise(new_totalstep = sum(new_steps, na.rm = T), totalstep = sum(steps, na.rm = T))
mean(temp$new_totalstep)
median(temp$new_totalstep)
```

- Do these values differ from the estimates from the first part of the assignment? YES

- What is the impact of imputing missing data on the estimates of the total daily number of steps?: Imputing missing data has led to an increase in both mean and median of the estimates of the total daily number of steps


#### 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
par(mfrow=c(2,1))
hist(temp$totalstep, xlab = "No.of steps each day", main = "Histogram of the total number of steps taken each day (without NAs)")
hist(temp$new_totalstep, xlab = "No.of steps each day", main = "Histogram of the total number of steps taken each day (imputing NAs)")
```

#### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(lattice)
newdf <- newdf %>% mutate(dayoftheweek = weekdays(date)) %>% 
        mutate(day_group = ifelse(dayoftheweek %in% c("Saturday", "Sunday"), "weekend", "weekday" )) %>% 
        mutate(day_group = as.factor(day_group))
temp <- newdf %>% group_by(interval, day_group) %>% summarise(average_step = mean(steps, na.rm = T))
xyplot(average_step ~ interval| day_group, data = temp, layout=c(1,2), type = "l",
       ylab = "Number of steps", xlab = "Interval")
```

 

